Assessing the Global COVID19 Impact on Air Transport with Open Data

### 2021-07-16 paper.Rmd
a few corrections and added some government reactions

For many years, many concerns of the global air traffic management community has been directed to the evident problem of imbalances between capacity and demand. The pressing, increasing demand for air transport registered in the last decade not only has already produced challenging delay management practices, but also fostered projections of even worse scenarios. EUROCONTROL [@statfor_growth_2018], for example, argued that delays in Europe could reach up to 20 minutes per flight in 2040, in stark contrast to the 12 minutes per flight, as registered in 2016.

@@

On March 11, 2020, the World Health Organisation has declared the novel coronavirus (COVID-19) outbreak a global pandemic.
The COVID-19 crisis posed a completely different, unexpected, and inverted challenge.
Communicable disease control resulted in massive restrictions on international and regional air traffic and passenger travel. 
Demand for air transport dropped as low as 90% of the previous "normal" in many places. Where the lack of capacity was previously the issue, now the lack of demand threatened the ATM system stability. 
In the financial perspective, airlines and airports had to deal with an unprecedented decrease in incomes. As a result, air navigation providers collected less fees for their services, due to significantly fewer flights. 
In the operational perspective, pilots and air traffic controllers practiced less. The problems and obstacles developed into many other dimensions.

@@

The unprecedented decline in air traffic demand resulted in severe financial strains on the air transport industry. 
The financial support or lack thereof for airlines and airports has been widely covered in the media. 
A variety of studies showed the interplay or consequences of the travel constraints. 
To date, lower attention was given to the inherent change in terms of air transport services
Hence, the current scenario is a proper moment to further investigate the concept of resilience.

@@

* provide some descriptive text about COVID and government reactions
* any more key milestones relevant for this study, e.g. US/EU stop, Lufthansa/BA/KLM/Air France stopping to operate to China as of late January

>>>>

For several years, many concerns of the global air traffic management community has been directed to the evident problem of imbalances between capacity and demand. The pressing, increasing demand for air transport registered in the last decade not only has already produced challenging delay management practices, but also fostered projections of even worse scenarios. EUROCONTROL [@statfor_growth_2018], for example, argued that delays in Europe could reach up to 20 minutes per flight in 2040, in stark contrast to the 12 minutes per flight, as registered in 2016.

@@

On March 11, 2020, the World Health Organisation has declared the novel coronavirus (COVID-19) outbreak a global pandemic.
The COVID-19 crisis posed a completely different, unexpected, and inverted challenge.
Communicable disease control resulted in massive restrictions on international and regional air traffic and passenger travel. 
Demand for air transport dropped as low as 90% of the previous "normal" in many places. Where the lack of capacity was previously the issue, now the lack of demand threatened the ATM system stability. 
In the financial perspective, airlines and airports had to deal with an unprecedented decrease in incomes. As a result, air navigation providers collected less fees for their services, due to significantly fewer flights. 
In the operational perspective, pilots and air traffic controllers practiced less. The problems and obstacles cascaded into many other dimensions.

@@

The unprecedented decline in air traffic demand resulted in severe financial strains on the air transport industry. 
The financial support or lack thereof for airlines and airports has been widely covered in the media. 
A variety of studies showed the interplay or consequences of the travel constraints. 
To date, lower attention was given to the inherent change in terms of air transport services.
Hence, the current scenario is a proper moment to further investigate the concept of resilience.

@@

* provide some descriptive text about COVID and government reactions (DONE BELOW!):

For example, in March 12, 2020, the United States established the suspension of most travels from the European Schengen area (BBC, 2020). A few days later, European Comission President Ursula von der Leyen announced a similar measure, proposing for EU Member States to apply restrictions on non-essential travel. By that time, several countries, such as Austria, Germany, and Poland had already implemented border check controls, while Portugal, Italy, and Spain have imposed lockdowns (Nielsen, 2020). At the same time, Australia imposed a severe arrival block on all non-citizens and non-residents at its airports (Worthington & Snape, 2020). In South America, Brazil, Chile, Argentina and other countries also followed the trend and closed their borders, totally or partially (France24, 2020). As a result, an unparalleled air traffic drop was seen worldwide.

BBC. (2020, March 12). Coronavirus: Trump suspends travel from Europe to US. https://www.bbc.com/news/world-us-canada-51846923

France24. (2020, March 16). Chile closes borders as Latam grounds flights over pandemic. https://www.france24.com/en/20200316-chile-closes-borders-as-latam-grounds-flights-over-pandemic

Nielsen, N. (2020, March 17). Commission proposes 'non-essential travel' EU entry ban. EU Observer. https://euobserver.com/coronavirus/147764

Worthington, B., & Snape, J. (2020, March 19). Australia blocks arrival of all non-citizens, non-residents in expanded coronavirus travel ban. ABC News. https://www.abc.net.au/news/2020-03-19/coronavirus-non-resident-travel-ban-australia/12071640

* any more key milestones relevant for this study, e.g. US/EU stop, Lufthansa/BA/KLM/Air France stopping to operate to China as of late January
